---
title: "brief1"
author: "..."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, message=FALSE, include=FALSE}
library(nationalparkcolors)
library(treemapify)
library(ggplot2)
library(dplyr)
library(kableExtra)
library(summarytools)
```

```{r, include=FALSE}
load(file = "~/GitHub/saetus/data_recode.RData")
```

```{r themeu, include=FALSE, message=FALSE, warning=FALSE}
#font_add_google(name = "Barlow", family = "foo")
#showtext::showtext_auto()

names(park_palettes)
park <- park_palette("Acadia", n = 3)
park2 <- park_palette("Redwoods", n = 5)
park3 <- park_palette("GeneralGrant", n = 7)
```

## Hardships

Since the onset of COVID-19 (March 2020), have you, or has anyone in your household experienced any of the following hardships: 
  
  Loss of employment income   
  Difficulty in paying utility bills (electricity, gas, cellphone, internet)   
  Difficulty in affording the kinds of food you need   
  Difficulty in paying next rent or mortgage payment   
  Difficulty in paying medical bill/health insurance 
  Delay getting/Do not get medical care or prescription medication

```{r, include=FALSE, eval=FALSE}
round(prop.table(table(data$bad, data$depen), 1)*100, 1)

round(prop.table(table(data$zipcode, data$bad), 1)*100, 1)

xtabs(formula = ~zipcode + bad + gender2, data = data, subset = data$gender2!="Other")
```
```{r}
kable(xtabs(formula = ~zipcode + gender2 + hardships_1, data = data, subset = data$gender2!="Other")) %>% 
  add_header_above(header = c("", "Loss of employment" = 3))

kable(xtabs(formula = ~zipcode + gender2 + hardships_2, data = data, subset = data$gender2!="Other")) %>% 
  add_header_above(header = c("", "Difficulty in paying utility bills" = 3))

kable(xtabs(formula = ~zipcode + gender2 + hardships_3, data = data, subset = data$gender2!="Other")) %>% 
  add_header_above(header = c("", "Difficulty in affording the food you need" = 3))

kable(xtabs(formula = ~zipcode + gender2 + hardships_4, data = data, subset = data$gender2!="Other")) %>% 
  add_header_above(header = c("", "Difficulty in paying next rent or mortgage" = 3))

kable(xtabs(formula = ~zipcode + gender2 + hardships_5, data = data, subset = data$gender2!="Other")) %>% 
  add_header_above(header = c("", "Difficulty in paying medical bill/health insurance" = 3))

kable(xtabs(formula = ~zipcode + gender2 + hardships_6, data = data, subset = data$gender2!="Other")) %>% 
  add_header_above(header = c("", "Delay getting medical care or prescription medication" = 3))
```


```{r, include=FALSE, message=FALSE}
data <- data %>% 
  mutate_at(vars(hardships_1:hardships_6), funs(recode(., "Yes" = as.numeric(1), "No" = as.numeric(0)))) %>% 
  mutate(hards = hardships_1+hardships_2+hardships_3+hardships_4+hardships_5+hardships_6)

data %>% 
  filter(!is.na(hards)) %>% 
  ggplot() + 
  geom_bar(aes(y = hards, fill = gender2)) + 
  facet_grid(~zipcode) +
  scale_fill_manual(values = park) +
  theme_classic() 

data %>% 
  filter(!is.na(hards), !is.na(educ)) %>% 
  ggplot() + 
  geom_bar(aes(y = hards, fill = educ)) + 
  facet_grid(~zipcode) +
  scale_fill_manual(values = park2) +
  theme_classic() 

data %>% 
  filter(!is.na(hards), !is.na(race_eth)) %>% 
  ggplot(aes(x = hards)) + 
  geom_bar(aes(y = ..prop..), stat = 'count') + 
  geom_text(aes(label = scales::percent(..prop.., accuracy = 0.1), y = ..prop..), 
            stat = "count", 
            vjust = .5, hjust = 0.2, size = 2.5, color = "black")+
  facet_grid(~race_eth) +
  scale_fill_manual(values = park3) +
  scale_y_continuous(labels = scales::percent)
```

```{r, include=FALSE, message=FALSE}
x <- data %>%
  filter(race_eth != "NH-Other") %>% 
  group_by(zipcode, race_eth) %>% 
  summarise(mean.hard = mean(hards, na.rm = T)) %>% 
  ungroup()

x2 <- data %>%
  group_by(zipcode, race_eth) %>%
  summarize(n = n(), av.hards =  mean(hards, na.rm = T)) %>% 
  mutate(pct = n/sum(n),
         lbl = scales::percent(pct, accuracy = 0.1))
```

```{r, include=FALSE, message=FALSE}
x %>%
  ggplot(aes(fill = race_eth, area = mean.hard,  label = race_eth)) +
  geom_treemap() + 
  geom_treemap_text(colour = "white", place = "centre") +
  scale_fill_manual(values = park3) +
  facet_grid(~zipcode) +
  theme(legend.position = "none")
```

```{r}
x2 %>% 
  ggplot(aes(x = race_eth, y = pct, fill = zipcode)) + 
  geom_bar(stat = "identity", position = "fill") +
  geom_text(aes(label = lbl), size = 3, color = "gray60", position = position_fill(vjust = .5)) +
  scale_y_continuous(breaks = seq(0, 1, .2), label = scales::percent) +
  scale_fill_manual(values = park) +
  labs(y = "Percent", 
       fill = "Race and Ethnicity",
       x = "zip code",
       caption = "2021 San Antonio Energy and Time Use Survey") +
  theme_minimal() +
  theme(legend.position = "top")
```

```{r, include=FALSE}
data %>% 
  filter(!is.na(gender2), !is.na(hards), gender2 != "Other") %>% 
  ggplot(aes(x = hards, 
           fill = gender2)) +
  geom_density(alpha = 0.5, color = "black") +
  labs(title = "Hardship distribution by zip code") +
  scale_fill_manual(values = park) +
  facet_grid(~zipcode) + 
  theme_classic()
```

```{r, include=FALSE}
data <- data %>% 
  mutate(bad = if_else(hards == 1, "One hardship", 
                       if_else(hards == 2, "Two hardships",
                               if_else(hards > 2, "More than two", NA_character_))))

data$bad <- ordered(data$bad, levels = c("One hardship", "Two hardships", "More than two"))

addmargins(table(data$educ, data$bad), margin = 2)
```

```{r, message=FALSE}
data %>% 
  filter(!is.na(bad), !is.na(educ)) %>% 
  group_by(zipcode, bad, educ) %>% 
  summarize(n = n()) %>% 
  mutate(pct = n/sum(n), 
         label = scales::percent(pct, accuracy = 0.1)) %>% 
  ggplot(aes(x = bad, y = pct, fill = educ)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_text(aes(label = label), size = 3, color = "black", position = position_fill(vjust = 0.5)) +
  scale_y_continuous(breaks = seq(0, 1, .15), label = scales::percent) +
  scale_fill_manual(values = park2) +
  facet_grid(~zipcode) +
  labs(y = NULL, 
       x = NULL,
       fill = NULL,
       caption = "2021 San Antonio Energy and Time Use Survey",
       title = "Distribution of hardships by educational level") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        plot.caption = element_text(size = 8), panel.grid = element_blank())
```

```{r, include=FALSE}
h1 <- data %>% 
  select(zipcode, hardships_1:hardships_6, age:bad) %>% 
  filter(bad == "One hardship") 

h2 <- data %>% 
  select(zipcode, hardships_1:hardships_6, age:bad) %>% 
  filter(bad == "Two hardships") 

h3 <- data %>% 
  select(zipcode, hardships_1:hardships_6, age:bad) %>% 
  filter(bad == "More than two") 
####################
h11 <- h1 %>% 
  group_by(zipcode) %>% 
  summarise_at(vars(hardships_1:hardships_6), funs(round(mean(., na.rm = T), 3)))
h12 <- h1 %>% 
  group_by(zipcode, gender2) %>% 
  summarise_at(vars(hardships_1:hardships_6), funs(round(mean(., na.rm = T), 3)))

h21 <- h2 %>% 
  group_by(zipcode) %>% 
  summarise_at(vars(hardships_1:hardships_6), funs(round(mean(., na.rm = T), 3)))
h22 <- h2 %>% 
  group_by(zipcode, gender2) %>% 
  summarise_at(vars(hardships_1:hardships_6), funs(round(mean(., na.rm = T), 3)))

h31 <- h3 %>% 
  group_by(zipcode) %>% 
  summarise_at(vars(hardships_1:hardships_6), funs(round(mean(., na.rm = T), 3)))
h32 <- h3 %>% 
  group_by(zipcode, gender2) %>% 
  summarise_at(vars(hardships_1:hardships_6), funs(round(mean(., na.rm = T), 3)))
################
h11 <- h11 %>%  
  tidyr::pivot_longer(cols = 2:7, names_to = "hard", values_to = "mean")
h12 <- h12 %>%  
  tidyr::pivot_longer(cols = 3:8, names_to = "hard", values_to = "mean")

h21 <- h21 %>%  
  tidyr::pivot_longer(cols = 2:7, names_to = "hard", values_to = "mean")
h22 <- h22 %>%  
  tidyr::pivot_longer(cols = 3:8, names_to = "hard", values_to = "mean")

h31 <- h31 %>%  
  tidyr::pivot_longer(cols = 2:7, names_to = "hard", values_to = "mean")
h32 <- h32 %>%  
  tidyr::pivot_longer(cols = 3:8, names_to = "hard", values_to = "mean")
###############
h11 <- h11 %>% 
  mutate(hard = case_when(hard %in% "hardships_1" ~ "Loss of employment",
                          hard %in% "hardships_2" ~ "Paying utility bills",
                          hard %in% "hardships_3" ~ "Affording food",
                          hard %in% "hardships_4" ~ "Paying next rent",
                          hard %in% "hardships_5" ~ "Paying medical bill",
                          hard %in% "hardships_6" ~ "Delay getting medical"))
h12 <- h12 %>% 
  mutate(hard = case_when(hard %in% "hardships_1" ~ "Loss of employment",
                          hard %in% "hardships_2" ~ "Paying utility bills",
                          hard %in% "hardships_3" ~ "Affording food",
                          hard %in% "hardships_4" ~ "Paying next rent",
                          hard %in% "hardships_5" ~ "Paying medical bill",
                          hard %in% "hardships_6" ~ "Delay getting medical"))

h21 <- h21 %>% 
  mutate(hard = case_when(hard %in% "hardships_1" ~ "Loss of employment",
                          hard %in% "hardships_2" ~ "Paying utility bills",
                          hard %in% "hardships_3" ~ "Affording food",
                          hard %in% "hardships_4" ~ "Paying next rent",
                          hard %in% "hardships_5" ~ "Paying medical bill",
                          hard %in% "hardships_6" ~ "Delay getting medical"))
h22 <- h22 %>% 
  mutate(hard = case_when(hard %in% "hardships_1" ~ "Loss of employment",
                          hard %in% "hardships_2" ~ "Paying utility bills",
                          hard %in% "hardships_3" ~ "Affording food",
                          hard %in% "hardships_4" ~ "Paying next rent",
                          hard %in% "hardships_5" ~ "Paying medical bill",
                          hard %in% "hardships_6" ~ "Delay getting medical"))

h31 <- h31 %>% 
  mutate(hard = case_when(hard %in% "hardships_1" ~ "Loss of employment",
                          hard %in% "hardships_2" ~ "Paying utility bills",
                          hard %in% "hardships_3" ~ "Affording food",
                          hard %in% "hardships_4" ~ "Paying next rent",
                          hard %in% "hardships_5" ~ "Paying medical bill",
                          hard %in% "hardships_6" ~ "Delay getting medical"))
h32 <- h32 %>% 
  mutate(hard = case_when(hard %in% "hardships_1" ~ "Loss of employment",
                          hard %in% "hardships_2" ~ "Paying utility bills",
                          hard %in% "hardships_3" ~ "Affording food",
                          hard %in% "hardships_4" ~ "Paying next rent",
                          hard %in% "hardships_5" ~ "Paying medical bill",
                          hard %in% "hardships_6" ~ "Delay getting medical"))
######################
h11$hard <- ordered(h11$hard, levels = c("Loss of employment", "Paying utility bills", "Affording food", 
                                         "Paying next rent", "Paying medical bill", "Delay getting medical"))
h12$hard <- ordered(h12$hard, levels = c("Loss of employment", "Paying utility bills", "Affording food", 
                                         "Paying next rent", "Paying medical bill", "Delay getting medical"))

h21$hard <- ordered(h21$hard, levels = c("Loss of employment", "Paying utility bills", "Affording food", 
                                         "Paying next rent", "Paying medical bill", "Delay getting medical"))
h22$hard <- ordered(h22$hard, levels = c("Loss of employment", "Paying utility bills", "Affording food", 
                                         "Paying next rent", "Paying medical bill", "Delay getting medical"))

h31$hard <- ordered(h31$hard, levels = c("Loss of employment", "Paying utility bills", "Affording food", 
                                         "Paying next rent", "Paying medical bill", "Delay getting medical"))
h32$hard <- ordered(h32$hard, levels = c("Loss of employment", "Paying utility bills", "Affording food", 
                                         "Paying next rent", "Paying medical bill", "Delay getting medical"))
```

```{r plots_1}
h11 %>% 
  mutate(label = scales::percent(mean, accuracy = 0.1)) %>% 
  ggplot(aes(x = hard, y = mean, fill = zipcode)) +
  geom_bar(stat = "identity", 
           position = "dodge", 
           color = "black") +
  geom_text(aes(y = mean, label = label), size = 3, color = "black", position = position_dodge(width = .9), vjust = -.5) +
  scale_fill_manual(values = park) +
  scale_y_continuous(breaks = seq(0, 1, .15), label = scales::percent) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL, 
       caption = "2021 San Antonio Energy and Time Use Survey", 
       title = "Hardship by zip code",
       subtitle = "Respondents with one hardship") +
  theme_classic() +
  theme(legend.position = c(.8, .8), legend.box.just = "right", legend.background = element_rect("gray95"), 
        plot.caption = element_text(size = 8), axis.text.x = element_text(size = 8), panel.background = element_rect("gray95"))

h21 %>% 
  mutate(label = scales::percent(mean, accuracy = 0.1)) %>% 
  ggplot(aes(x = hard, y = mean, fill = zipcode)) +
  geom_bar(stat = "identity", 
           position = "dodge", 
           color = "black") +
  geom_text(aes(y = mean, label = label), size = 3, color = "black", position = position_dodge(width = .9), vjust = -.5) +
  scale_fill_manual(values = park) +
  scale_y_continuous(breaks = seq(0, 1, .15), label = scales::percent) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL, 
       caption = "2021 San Antonio Energy and Time Use Survey", 
       title = "Hardship by zip code",
       subtitle = "Respondents with two hardships") +
  theme_classic() +
  theme(legend.position = c(.8, .8), legend.box.just = "right", legend.background = element_rect("gray95"), 
        plot.caption = element_text(size = 8), axis.text.x = element_text(size = 8), panel.background = element_rect("gray95"))

h31 %>% 
  mutate(label = scales::percent(mean, accuracy = 0.1)) %>% 
  ggplot(aes(x = hard, y = mean, fill = zipcode)) +
  geom_bar(stat = "identity", 
           position = "dodge", 
           color = "black") +
  geom_text(aes(y = mean, label = label), size = 3, color = "black", position = position_dodge(width = .9), vjust = -.5) +
  scale_fill_manual(values = park) +
  scale_y_continuous(breaks = seq(0, 1, .15), label = scales::percent) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL, 
       caption = "2021 San Antonio Energy and Time Use Survey", 
       title = "Hardship by zip code",
       subtitle = "Respondents with more than two hardship") +
  theme_classic() +
  theme(legend.position = c(.8, .8), legend.box.just = "right", legend.background = element_rect("gray95"), 
        plot.caption = element_text(size = 8), axis.text.x = element_text(size = 8), panel.background = element_rect("gray95"))
```

```{r plots_2}
h12 %>% 
  filter(gender2 != "Other") %>% 
  mutate(label = scales::percent(mean, accuracy = 0.1)) %>% 
  ggplot(aes(x = hard, y = mean, fill = gender2)) +
  geom_bar(stat = "identity", 
           position = "dodge", 
           color = "black") +
  geom_text(aes(y = mean, label = label), size = 3, color = "black", position = position_dodge(width = .9), vjust = -.5) +
  facet_grid(~zipcode) +
  scale_fill_manual(values = park) +
  scale_y_continuous(breaks = seq(0, 1, .15), label = scales::percent) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL, 
       caption = "2021 San Antonio Energy and Time Use Survey", 
       title = "Hardship by zip code",
       subtitle = "Respondents with one hardship") +
  theme_classic() +
  theme(legend.position = c(.8, .8), legend.box.just = "right", legend.background = element_rect("gray95"), 
        plot.caption = element_text(size = 8), axis.text.x = element_text(size = 8), panel.background = element_rect("gray95"))

h22 %>% 
  filter(gender2 != "Other") %>% 
  mutate(label = scales::percent(mean, accuracy = 0.1)) %>% 
  ggplot(aes(x = hard, y = mean, fill = gender2)) +
  geom_bar(stat = "identity", 
           position = "dodge", 
           color = "black") +
  geom_text(aes(y = mean, label = label), size = 3, color = "black", position = position_dodge(width = .9), vjust = -.5) +
  facet_grid(~zipcode) +
  scale_fill_manual(values = park) +
  scale_y_continuous(breaks = seq(0, 1, .15), label = scales::percent) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL, 
       caption = "2021 San Antonio Energy and Time Use Survey", 
       title = "Hardship by zip code",
       subtitle = "Respondents with one hardship") +
  theme_classic() +
  theme(legend.position = c(.8, .8), legend.box.just = "right", legend.background = element_rect("gray95"), 
        plot.caption = element_text(size = 8), axis.text.x = element_text(size = 8), panel.background = element_rect("gray95"))

h32 %>% 
  filter(gender2 != "Other") %>% 
  mutate(label = scales::percent(mean, accuracy = 0.1)) %>% 
  ggplot(aes(x = hard, y = mean, fill = gender2)) +
  geom_bar(stat = "identity", 
           position = "dodge", 
           color = "black") +
  geom_text(aes(y = mean, label = label), size = 3, color = "black", position = position_dodge(width = .9), vjust = -.5) +
  facet_grid(~zipcode) +
  scale_fill_manual(values = park) +
  scale_y_continuous(breaks = seq(0, 1, .15), label = scales::percent) +
  labs(x = NULL, 
       y = NULL, 
       fill = NULL, 
       caption = "2021 San Antonio Energy and Time Use Survey", 
       title = "Hardship by zip code",
       subtitle = "Respondents with one hardship") +
  theme_classic() +
  theme(legend.position = c(.9, .85), legend.box.just = "right", legend.background = element_rect("gray95"), 
        plot.caption = element_text(size = 8), axis.text.x = element_text(size = 8), panel.background = element_rect("gray95"))

```



```{r}
view(dfSummary(h1, plain.ascii = F, graph.magnif = .75, labels.col = T, max.string.width = 15), method = "render")
```
```{r}
view(dfSummary(h2, plain.ascii = F, graph.magnif = .75, labels.col = T, max.string.width = 15), method = "render")
```

```{r}
view(dfSummary(h3, plain.ascii = F, graph.magnif = .75, labels.col = T, max.string.width = 15), method = "render")
```


























