---
title: "San Antonio Energy and Time Use Survey"
author: "..."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(readxl)
library(dplyr)
library(ggplot2)
library(plotly)
library(patchwork)
library(showtext)
library(kableExtra)
library(summarytools)
library(gtsummary)
```

```{r load, include=FALSE}
# From Qualtrics we get the data as .csv with default options: Recode seen but unanswered questions as -99. Recode seen but unanswered multi-value fields as 0.
# Change path
san <- read_excel("~/GitHub/saetus/San Antonio Energy and Time Use Survey (03.10).xlsx", col_names = T)
# delete responses before 22.02. These were responses from us. Delete some defaults columns, only have NA values.
san <- san[-c(1:9), - c(10:13)] 
```

```{r NAs, include=FALSE}
# 174 variables when uploaded.  175 after this point.
# disagree; other_major_imp; other_impairment_hsh; other_eip

san <- san %>% 
  # I'm subtracting the column (rewrite_email) since nobody answer the question
  mutate(number_missing_value = rowSums(is.na(san))-1,
         # 174 total possible responses
         nas_pct = round((number_missing_value/174*100), 1))

### NAs
summary(san$number_missing_value)

### by Peter
san <- san %>%
  mutate(duration = round(as.numeric(`Duration (in seconds)`)/60, 1),
         upper_out = quantile(duration, probs = 0.975, na.rm = T),
         upper_out2 = quantile(duration, probs = 0.95, na.rm = T), 
         lower_out = quantile(duration, probs = 0.025, na.rm = T),
         lower_out2 = quantile(duration, probs = 0.05, na.rm = T),
         na_low = quantile(nas_pct, probs = 0.10),
         na_upp = quantile(nas_pct, probs = 0.90),
         duration_min_out = if_else(duration <= lower_out, "Spend too little time", 
                                           if_else(duration >= upper_out, "Spend too much time", "Normal")),
         duration_min_out2 = if_else(duration <= lower_out2, "Below 5% quantile", 
                                           if_else(duration >= upper_out2, "Above 95% quantile", "Normal")),
         nas_out = if_else(nas_pct <= na_low, "Below 10% quantile",
                           if_else(nas_pct >= na_upp, "Above 90% quantile", "Normal")))

san$duration_min_out <- ordered(san$duration_min_out, levels = c("Spend too much time", "Normal", "Spend too little time"))
san$duration_min_out2 <- ordered(san$duration_min_out2, levels = c("Above 95% quantile", "Normal", "Below 5% quantile"))
san$nas_out <- ordered(san$nas_out, levels = c("Above 90% quantile", "Normal", "Below 10% quantile"))


san <- san %>% 
  mutate(odd_r = if_else(duration_min_out != "Normal" & nas_out == "Above 90% quantile", "Odd response", "Normal"),
         odd_r2 = if_else(duration_min_out2 != "Normal" & nas_out == "Above 90% quantile", "Odd response", "Normal"))
```

```{r recode, include=FALSE}
san <- san %>% 
  # consent Agree and double check on survey participation. Keep only zip codes 78202 or 78230.
  #filter(consent == "1" | disagree == "1", zipcode != "3") %>% 
  mutate(yrbr = ifelse(as.numeric(year) >= 1, 2000 - as.numeric(year), NA), # year of birth
         age = 2021 - yrbr, # age
         zip = if_else(zipcode == "1", "Elm Creek", "Jefferson Heights"), # recode zip codes
         hisp = if_else(hispanic == "1", "Not Hispanic", if_else(hispanic == "2", "Hispanic", "Don't know")), # Hispanic origin
         race2 = case_when(race %in% "1" ~ "White", # recode race
                           race %in% "2" ~ "Black", 
                           race %in% "3" ~ "Asian", 
                           race %in% "4" ~ "AIAN", # American Indian or Alaskan Native
                           race %in% "5" ~ "NHPI", # Native Hawaiian of Other Pacific Islander
                           race %in% c("6", "7", "8") ~ "Other"), # Two or more races; Unknown|Don't know; Prefer not to answer
         race_eth = if_else(hisp != "Hispanic" & race2 == "White", "NH-White", 
                            if_else(hisp != "Hispanic" & race2 == "Black", "NH-Black", 
                                    if_else(hisp != "Hispanic" & race2 == "Asian", "NH-Asian", 
                                            if_else(hisp != "Hispanic" & race2 == "AIAN", "NH-AIAN",
                                                    if_else(hisp != "Hispanic" & race2 == "NHPI", "NH-NHPI",
                                                            if_else(hisp != "Hispanic" & race2 == "Other", "NH-Other", "Hispanic")))))),
         gender = case_when(gender %in% "1" ~ "Female", 
                            gender %in% "2" ~ "Male", 
                            gender %in% c("3", "4", "5", "6", "7") ~ "Other"),
         marital_st = case_when(marital %in% "1" ~ "Single",
                                marital %in% c("2", "3") ~ "In a union", # married/living in a marriage-like relationship
                                marital %in% c("4", "5", "6") ~ "Not in a union"), # widowed, separated, divorced
         educ = case_when(education %in% "1" ~ "Less than HS", 
                          education %in% c("2", "3") ~ "High School", # some high school, high school graduate or equivalent 
                          education %in% c("4", "5", "6") ~ "College", # some college, AA, BA, BS
                          education %in% "7" ~ "Graduate",
                          education %in% "8" ~ "Don't know"),
         health = case_when(health %in% "1" ~ "Excellent", health %in% "2" ~ "Very good", health %in% "3" ~ "Good", 
                            health %in% "4" ~ "Fair", health %in% "5" ~ "Poor"))

san$race_eth <- ordered(san$race_eth, levels = c("NH-White", "NH-Black", "NH-Asian", "NH-AIAN", "NH-NHPI", "NH-Other", "Hispanic"))
san$health <- ordered(san$health, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))

# Total responses 849 - not including responses "no Finished"
# Total responses 857 - including responses "no Finished"
# length(unique(san$email)) 
```

1) Locate "odd responses" based on duration and number of NA in each response
2) Duration: a) Keep responses in (2.5%, 97.5%), b) responses in (5%, 95%)
3) NAs: a) Responses in (10%, 90%)

```{r, message=FALSE, warning=FALSE}
san %>% 
  #filter(!is.na(zipcode)) %>% 
  ggplot(aes(x = zipcode, fill = duration_min_out)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(0.5), vjust = 0, size = 2, color = "black") +
  scale_fill_brewer(type = "div", palette = 'Spectral') +
  #facet_grid(~ zipcode) +
  ggtitle("Duration between 0.7 min and 97.3 min = NORMAL (95% of responses)") +
  xlab("Zip code") + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, size = 9), 
        panel.background = element_rect(fill = "gray95"))

san %>% 
  #filter(!is.na(zipcode)) %>% 
  ggplot(aes(x = zipcode, fill = duration_min_out2)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(0.5), vjust = 0, size = 2, color = "black") +
  scale_fill_brewer(type = "div", palette = 'Spectral') +
  #facet_grid(~ zipcode) +
  ggtitle("Duration between 1.36 min and 71.2 min = NORMAL (90% of reponses)") +
  xlab("Zip code") + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, size = 9), 
        panel.background = element_rect(fill = "gray95"))

san %>% 
  #filter(!is.na(zipcode)) %>% 
  ggplot(aes(x = zipcode, fill = nas_out)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(0.5), vjust = 0, size = 2, color = "black") +
  scale_fill_brewer(type = "div", palette = 'Spectral') +
  #facet_grid(~ zipcode) +
  ggtitle("Number of NAs between 6 and 91 = NORMAL") +
  xlab("Zip code") + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, size = 9), 
        panel.background = element_rect(fill = "gray95"))



plot_ly(data = san, x = ~nas_pct, y = ~duration, color = ~duration_min_out2) %>% 
  layout(title = "Duration (min) by NAs by response (%)", 
         yaxis = list(title = "Duration (min)"),
         xaxis = list(title = "Percentage of NA values"))

san %>% 
  filter(!is.na(zipcode)) %>% 
  ggplot() +
  geom_boxplot(aes(y = nas_pct, x = duration_min_out)) +
  scale_fill_brewer(type = "div", palette = 'Spectral') +
  ggtitle("NA dist. by duration in minutes") +
  xlab("Duration between 0.7 min and 97.3 min = NORMAL (95% of responses)") + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 9), 
        panel.background = element_rect(fill = "gray95"))

san %>% 
  filter(!is.na(zipcode)) %>% 
  ggplot() +
  geom_boxplot(aes(y = nas_pct, x = duration_min_out, fill = as.factor(zipcode))) +
  scale_fill_brewer(type = "div", palette = 'Spectral') +
  ggtitle("NA dist. by duration in minutes by zip code") +
  xlab("Duration between 0.7 min and 97.3 min = NORMAL (95% of responses)") + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 9), 
        panel.background = element_rect(fill = "gray95"))

san %>% 
  filter(!is.na(zipcode), duration < 900) %>% 
  ggplot() +
  geom_boxplot(aes(y = duration, fill = as.factor(zipcode), x = nas_out)) +
  scale_fill_brewer(type = "div", palette = 'Spectral') +
  ggtitle("Duration (min) by zip code and NA distribution") +
  xlab("Zip code") + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 9), 
        panel.background = element_rect(fill = "gray95"))
```

```{r, include=FALSE, message=FALSE, warning=FALSE}
# average duration by zipcode/race
tab <- san %>% 
  group_by(zip, race_eth) %>% 
  dplyr::summarise(av.duration = round(mean(duration, na.rm = T), 1), sd.duration = round(sd(duration, na.rm = T), 1))

tab.z <- san %>% 
  group_by(zip) %>% 
  dplyr::summarise(av.duration = round(mean(duration, na.rm = T), 1), sd.duration = round(sd(duration, na.rm = T), 1))

kable(tab, format = "html")

kable(tab.z, format = "html")

#kableExtra::kable(table(san$zip, san$gender))
#kableExtra::kable(round(prop.table(table(san$zip, san$gender), 2)*100, 1))
```

```{r recode_health, include=FALSE, eval=FALSE}
test <- san %>% 
  #filter(!is.na(age)) %>% # what to do with the responses not finished
  mutate(ageg = case_when(age %in% min(age):29 ~ "20-29",
                                                 age %in% 30:39 ~ "30-39",
                                                 age %in% 40:49 ~ "40-49",
                                                 age %in% 50:59 ~ "50-59",
                                                 age %in% 60:max(age) ~ "60+"))


min(san$age)

san <- san %>% 
  #filter(Finished == "1") %>% # what to do with the responses not finished
  mutate(ageg = case_when(age %in% min(age):29 ~ "20-29",
                          age %in% 30:39 ~ "30-39",
                          age %in% 40:49 ~ "40-49",
                          age %in% 50:59 ~ "50-59",
                          age %in% 60:max(age) ~ "60+", FALSE ~ NA_character_),
         anxious = if_else(mental_health_anxious == "1", "Not at all",
                           if_else(mental_health_anxious == "2", "Several days", 
                                   if_else(mental_health_anxious == "3", "More than half",
                                           if_else(mental_health_anxious == "4", "Nearly every day", 
                                                   if_else(mental_health_anxious != c("1", "2", "3", "4"), "NA", mental_health_anxious))))),
         worry = if_else(mental_health_worry == "1", "Not at all",
                           if_else(mental_health_worry == "2", "Several days", 
                                   if_else(mental_health_worry == "3", "More than half",
                                           if_else(mental_health_worry == "4", "Nearly every day", 
                                                   if_else(mental_health_worry != c("1", "2", "3", "4"), "NA", mental_health_worry))))),
         interest = if_else(mental_health_interest == "1", "Not at all",
                           if_else(mental_health_interest == "2", "Several days", 
                                   if_else(mental_health_interest == "3", "More than half",
                                           if_else(mental_health_interest == "4", "Nearly every day", 
                                                   if_else(mental_health_interest != c("1", "2", "3", "4"), "NA", mental_health_interest))))),
         down = if_else(mental_health_down == "1", "Not at all",
                           if_else(mental_health_down == "2", "Several days", 
                                   if_else(mental_health_down == "3", "More than half",
                                           if_else(mental_health_down == "4", "Nearly every day", 
                                                   if_else(mental_health_down != c("1", "2", "3", "4"), "NA", mental_health_down))))),
         pre_c19_ph = if_else(pre_phy_health == "1", "Worse", 
                              if_else(pre_phy_health == "2", "Same", 
                                      if_else(pre_phy_health == "3", "Better", pre_phy_health))),
         pre_c19_mh = if_else(pre_mental_health == "1", "Worse", 
                              if_else(pre_mental_health == "2", "Same", 
                                      if_else(pre_mental_health == "3", "Better", pre_mental_health))))

san$anxious <- ordered(san$anxious, levels = c("Not at all", "Several days", "More than half", "Nearly every day"))
san$worry <- ordered(san$worry, levels = c("Not at all", "Several days", "More than half", "Nearly every day"))
san$interest <- ordered(san$interest, levels = c("Not at all", "Several days", "More than half", "Nearly every day"))
san$down <- ordered(san$down, levels = c("Not at all", "Several days", "More than half", "Nearly every day"))
san$pre_c19_ph <- ordered(san$pre_c19_ph, levels = c("Worse", "Same", "Better"))
san$pre_c19_mh <- ordered(san$pre_c19_mh, levels = c("Worse", "Same", "Better"))
```

```{r recode_employment, include=FALSE, eval=FALSE}
san <- san %>% 
  mutate(front = if_else(frontline == "1", "Yes", 
                         if_else(frontline == "2", "No", "NA")),
         work_hrs = as.numeric(hrs_work_3),
         hsh_size = as.numeric(num_hsh),
         hsh_child = as.numeric(num_children),
         hsh_retire = as.numeric(num_retired))
```

```{r health_status_plot, message=FALSE, warning=FALSE, eval=FALSE}
load("~/GitHub/san/sa_survey.RData")

san %>% 
  filter(health != "NA") %>% 
  ggplot(aes(x = race_eth, fill = health)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(0.5), vjust = 0, size = 2) +
  scale_fill_viridis_d(option = "E", begin = .2, end = .8) +
  facet_grid(~zip) +
  ggtitle("Responses by zip code and mental health status", 
          subtitle = "How often have you been bothered by the feeling nervous, anxious, or on edge") +
  xlab(NULL) + 
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, size = 9))

san %>% 
  filter(anxious != "NA", worry != "NA", interest != "NA", down!= "NA") %>% 
  ggplot(aes(x = race_eth, fill = anxious)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(0.5), vjust = 0, size = 2) +
  scale_fill_viridis_d(option = "E", begin = .2, end = .8) +
  facet_grid(~zip) +
  ggtitle("Responses by zip code and mental health status", 
          subtitle = "How often have you been bothered by the feeling nervous, anxious, or on edge") +
  xlab(NULL) + 
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, size = 9))

san %>% 
  filter(anxious != "NA", worry != "NA", interest != "NA", down!= "NA", !is.na(race_eth)) %>% 
  ggplot(aes(x = race_eth, fill = worry)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(0.5), vjust = 0, size = 2) +
  scale_fill_viridis_d(option = "E", begin = .2, end = .8) +
  facet_grid(~ zip) +
  ggtitle("Responses by zip code and mental health status", 
          subtitle = "How often have you been bothered by not being able to stop or control worrying") +
  xlab(NULL) + 
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, size = 9))

san %>% 
  filter(anxious != "NA", worry != "NA", interest != "NA", down!= "NA", !is.na(race_eth)) %>% 
  ggplot(aes(x = race_eth, fill = interest)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(0.5), vjust = 0, size = 2) +
  scale_fill_viridis_d(option = "E", begin = .2, end = .8) +
  facet_grid(~ zip) +
  ggtitle("Responses by zip code and mental health status", 
          subtitle = "How often have you been bothered by having little interest or pleasure in doing things") +
  xlab(NULL) + 
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, size = 9))

san %>% 
  filter(anxious != "NA", worry != "NA", interest != "NA", down!= "NA", !is.na(race_eth)) %>% 
  ggplot(aes(x = race_eth, fill = down)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(.5), vjust = 0, size = 2) +
  scale_fill_viridis_d(option = "E", begin = .2, end = .8) +
  facet_grid(~ zip) +
  ggtitle("Responses by zip code and mental health status", 
          subtitle = "How often have you been bothered by feeling down, depressed, or hopeless") +
  xlab(NULL) + 
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, size = 9))

san %>% 
  filter(pre_c19_ph != "NA", pre_c19_mh != "NA", !is.na(race_eth)) %>% 
  ggplot(aes(x = race_eth, fill = pre_c19_ph)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(.5), vjust = 0, size = 2) +
  scale_fill_viridis_d(option = "D", begin = .2, end = .8) +
  facet_grid(~ zip) +
  ggtitle("Responses by zip code and physical health prior to COVID-19") +
  xlab(NULL) + 
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, size = 9))

san %>% 
  filter(pre_c19_ph != "NA", pre_c19_mh != "NA", !is.na(race_eth)) %>% 
  ggplot(aes(x = race_eth, fill = pre_c19_mh)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(.5), vjust = 0, size = 2) +
  scale_fill_viridis_d(option = "D", begin = .2, end = .8) +
  facet_grid(~ zip) +
  ggtitle("Responses by zip code and mental health prior to COVID-19") +
  xlab(NULL) + 
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, size = 9))
```

```{r frontline_plot, eval=FALSE}
san %>% 
  filter(front != "NA") %>% 
  ggplot(aes(x = race_eth, fill = front)) +
  geom_bar() +
  geom_text(aes(label = ..count..), stat = "count", position=position_stack(0.5), vjust = 0, size = 2, color = "white") +
  scale_fill_viridis_d(option = "B", begin = .2, end = .8) +
  facet_grid(~ zip) +
  ggtitle("Responses by zip code and identification as frontline worker") +
  xlab(NULL) + 
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, size = 9))

san %>% 
  filter(front != "NA",  gender != "NA", work_hrs > 0 ) %>% 
  ggplot() +
  geom_col(aes(x = work_hrs, y = race_eth, color = gender, fill = gender), position = "dodge", width = .5) +
  facet_grid(~ zip) +
  scale_color_viridis_d(option = "E", begin = .2, end = .8) +
  scale_fill_viridis_d(option = "E", begin = .2, end = .8) +
  ggtitle("Responses by zip code and working hours per week") +
  xlab("Hours per week") + 
  ylab(NULL) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 90, size = 9))
```

```{r rep_emails, include=FALSE, eval=FALSE}
### Replicated email addresses 
unique_emails <- unique(san$email)
rep_emails <- c("AdrianaHudsonwhp@yahoo.com", "AliRushwynf@yahoo.com", "altenwerthvbwht42@gmail.com", "ArdenHubbardvqk@yahoo.com",
               "ArnoldRatigmw@gmail.com", "AthenaHerringjuw@yahoo.com", "augustusyzlk19@gmail.com", "charlesfpsk94@gmail.com",
               "choeaqzn14@gmail.com", "choeaqzn14@gmail.com", "ChristinaEdwardsaocv@yahoo.com", "deboeremh85@gmail.com", 
               "DianaRiosavb@yahoo.com", "dooleyqigx99@gmail.com", "EliGrahamfsxa@yahoo.com", "fochttdlfv47@gmail.com",
               "harrisckzi97@gmail.com", "hubbardpaul564@gmail.com", "IngridBuckleyuqgg@yahoo.com", "JaelynPettyvah@yahoo.com",
               "JeffreyRojaseeko@yahoo.com", "LatoyaParkermkiu@yahoo.com", "MalcomRyantom@yahoo.com", "NanetteWrighthys@yahoo.com",
               "PedroDawsonsifc@yahoo.com", "PhillipBuckleygmmm@yahoo.com", "RafaelManningygoo@yahoo.com", "rodggersmmaison@gmail.com",
               "SheaMillerkcbe@yahoo.com", "tameracaqcx43@gmail.com", "TraciRivasrqy@yahoo.com", "TyrellCowanjqm@yahoo.com", 
               "ValerieLevyiuaz@yahoo.com", "wheelerjoselyn@gmail.com")
san <- san %>% 
  mutate(rep_email = if_else(email == lag(email, n = 1, order_by = email), "rep_email", "first_entry"))

san_re <- san %>% 
  mutate(rep_email = if_else(email == lag(email, n = 1, order_by = email), "rep_email", "first_entry")) %>% 
  filter(email %in% rep_emails) %>% 
  arrange(email)

save(san, file = "~/GitHub/san/sa_survey.RData")
save(san_re, file = "~/GitHub/san/sa_survey_rep_emails.RData")

write.csv(x = unique_emails, file = "~/GitHub/san/unique_emails.csv")
write.csv(x = rep_emails, file = "~/GitHub/san/repeated_emails.csv")

write.csv(x = san, file = "~/GitHub/san/sa_energy_survey.csv", row.names = F)
write.csv(x = san_re, file = "~/GitHub/san/sa_energy_survey_rep_emails.csv", row.names = F)
```

```{r emails, include=FALSE, eval=FALSE}
### EMAILS
#### IDs
emails <- san %>% 
  mutate(rep_email = if_else(email == lag(email, n = 1, order_by = email), "rep_email", "first_entry"),
         mean_duration = round(mean(duration, na.rm = T), 1),
         median_duration = median(duration, na.rm = T),
         nas = rowSums(is.na(.)) - 4,
         na_pct = round(nas/161 * 100, 1)) %>% 
  filter(rep_email == "first_entry" & zip != 3) %>% 
  select(ResponseId, email, zip, duration, mean_duration, median_duration, nas, na_pct) %>%  
  arrange(duration, na_pct)


write.csv(x = emails, file = "~/GitHub/saetus/emails.csv", row.names = F)
```

```{r, include=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
font_add_google(name = "Barlow", family = "foo")
showtext::showtext_auto()

themeu <- theme(plot.title = element_text(size = 22, face = "bold"), 
                plot.subtitle = element_text(size = 11),
                legend.position = "none", 
                legend.title = element_blank(), 
                axis.title = element_text(size = 17), 
                axis.text = element_text(size = 17), 
                axis.text.y = element_text(size = 17, color = "black"),
                plot.caption = element_text(size = 12), 
                panel.grid = element_blank(),
                strip.background = element_rect(fill = NULL, linetype = "solid"), 
                strip.text = element_text(color = "white", size = 18, face = "bold"))
```

```{r, include=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
san %>% 
  group_by(race_eth, zip, gender) %>% 
  dplyr::summarise(poorh = round(mean(as.numeric(days_poor_health), na.rm = T), 1)) %>% 
  ungroup() %>% 
  select(race_eth, zip, gender, poorh) %>% 
  filter(gender != "Other", race_eth!= "NH-Other") %>% 
  mutate(label = poorh) %>% 
  ggplot(aes(x = poorh, y = race_eth)) +
  geom_line(color = "gray50", size = 1) +
  geom_point(aes(fill = as.factor(gender), color = as.factor(gender)), size = 4, shape = 19) +
  facet_grid(~zip) +
  geom_text(aes(label = label, x = poorh), nudge_y = 0.25, size = 4.5) +
  geom_text(aes(label = gender, x = poorh), nudge_y = -0.25, size = 4.5) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual") +
  labs(title = "Days of poor health | Women and Men by race and ethnicity",
    x = "Average number of days with poor physical or mental health in January 2021",
    y = NULL, 
    caption = "Source: 2021 San Antonio Energy and Time Use Survey") +
  theme_linedraw() +
  themeu
  

san %>% 
  group_by(educ, zip, gender) %>% 
  dplyr::summarise(poorh = round(mean(as.numeric(days_poor_health), na.rm = T), 1)) %>% 
  ungroup() %>% 
  select(educ, zip, gender, poorh) %>% 
  filter(gender != "Other", educ != "NA") %>% 
  mutate(label = poorh) %>% 
  ggplot(aes(x = poorh, y = educ)) +
  geom_line(color = "gray50", size = 1) +
  geom_point(aes(fill = as.factor(gender), color = as.factor(gender)), size = 4, shape = 19) +
  facet_grid(~zip) +
  geom_text(aes(label = label, x = poorh), nudge_y = 0.25, size = 4.5) +
  geom_text(aes(label = gender, x = poorh), nudge_y = -0.25, size = 4.5) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual") +
  labs(title = "Days of poor health | Women and Men by education level",
    x = "Average number of days with poor physical or mental health in January 2021", 
    y = NULL, 
    caption = "Source: 2021 San Antonio Energy and Time Use Survey") +
  theme_linedraw() +
  themeu

san %>% 
  group_by(front, zip, gender) %>% 
  dplyr::summarise(poorh = round(mean(as.numeric(days_poor_health), na.rm = T), 1)) %>% 
  ungroup() %>% 
  select(front, zip, gender, poorh) %>% 
  filter(gender != "Other", front != "NA") %>% 
  mutate(label = poorh) %>% 
  ggplot(aes(x = poorh, y = front)) +
  geom_line(color = "gray50", size = 1) +
  geom_point(aes(fill = as.factor(gender), color = as.factor(gender)), size = 4, shape = 19) +
  facet_grid(~zip) +
  geom_text(aes(label = label, x = poorh), nudge_y = 0.25, size = 4.5) +
  geom_text(aes(label = gender, x = poorh), nudge_y = -0.25, size = 4.5) +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  scale_fill_brewer(type = "qual") +
  labs(title = "Days of poor health | Women and Men by frontline worker status",
    x = "Average number of days with poor physical or mental health in January 2021", 
    y = NULL, 
    caption = "Source: 2021 San Antonio Energy and Time Use Survey") +
  theme_linedraw() +
  themeu
```

## Part I.

```{r dfSummary, message=FALSE, warning=FALSE}
view(dfSummary(san[, 5:173], plain.ascii = F, graph.magnif = .75, labels.col = T, max.string.width = 15), method = "render")
```

## Part II.

### Which responses to keep?

1) Consider odd response those outside the "normal" range of duration (1.36, 71) AND above 91 NA values 
2) Summary table: responses that Agree to participate, valid zip code (78202 or 78230), not an odd response

```{r, message=FALSE, warning=FALSE}
### Duration and number of NAs
kable(table(san$nas_out, san$duration_min_out), caption = "NAs (%) vs Duration (min)") %>% 
  kable_classic(full_width = F) %>% 
  footnote(general = "NAs in (10%, 90%); duration in (2.5%, 97.5%)")

kable(table(san$nas_out, san$duration_min_out2), caption = "NAs (%) vs Duration (min)") %>% 
  kable_classic(full_width = F) %>% 
  footnote(general = "NAs in (10%, 90%); duration in (5%, 95%)")  



kable(table(san$odd_r2), caption = "Odd responses") %>% 
  kable_classic(full_width = F)


sank <- san %>% 
  # consent Agree and double check on survey participation. Keep only zip codes 78202 or 78230. And keep only the first answer for unique emails
  filter(consent == "1" | disagree == "1", zipcode != 3, odd_r2 == "Normal") %>% 
  mutate(zipcode = if_else(zipcode == "1", "Elm Creek", 
                           if_else(zipcode == "2", "Jefferson Heights", NA_character_)),
         rep_email = if_else(email == lag(email, n = 1, order_by = email), "rep_email", "first_entry")) %>% 
  filter(rep_email == "first_entry")

#view(dfSummary(sank[, 5:173], plain.ascii = F, graph.magnif = .75, labels.col = T, max.string.width = 15), method = "render")
```

### Section IV. Background Characteristics by Zip code

```{r section_iv, message=FALSE, warning=FALSE}
sank %>%
<<<<<<< HEAD
  select(16, 146, 148:171, 175, 178:189) %>% 
=======
  select(146, 148:171, 175, 178:189) %>% 
>>>>>>> f1fcb7a0f07dfade34b2992b98c3fb2571364b8b
  tbl_summary(by = zipcode, 
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(1, 1))) %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**variable**") %>% 
  modify_caption("**Section IV. Background characteristics (N = {N})**") %>% 
  bold_labels()
```

### Section Ia. Time Use by Zip code

```{r section_ia, message=FALSE, warning=FALSE}
sank %>%
  select(16:80, 175, 178:189) %>% 
  tbl_summary(by = zipcode, 
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(1, 1))) %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**variable**") %>% 
  modify_caption("**Section Ia. Time Use (N = {N})**") %>% 
  bold_labels()
```


### Section Ib. Energy Use by Zip code

```{r section_ib, message=FALSE, warning=FALSE}
sank %>%
  select(16, 81:87, 175, 178:189) %>% 
  tbl_summary(by = zipcode, 
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(1, 1))) %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**variable**") %>% 
  modify_caption("**Section Ib. Energy Use (N = {N})**") %>% 
  bold_labels()
```

### Section II. Health Conditions by Zip code

```{r section_ii, message=FALSE, warning=FALSE}
sank %>%
  select(16, 88:101, 103, 105:119, 175, 178:189) %>% 
  tbl_summary(by = zipcode, 
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(1, 1))) %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**variable**") %>% 
  modify_caption("**Section II. Health Conditions (N = {N})**") %>% 
  bold_labels()
```

### Section III. Employment by Zip code

```{r section_iii, message=FALSE, warning=FALSE}
sank %>%
  select(16, 121:143, 175, 178:189) %>% 
  tbl_summary(by = zipcode, 
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                               all_categorical() ~ "{n} ({p}%)"),
              digits = list(all_continuous() ~ c(1, 1))) %>% 
  add_n() %>% 
  add_p() %>% 
  modify_header(label = "**variable**") %>% 
  modify_caption("**Section III. Employment (N = {N})**") %>% 
  bold_labels()
```















